---
title: "Analyse  the strength of associations between binary or binomial distributed outcomes and exposure variables"
#author: "David"
#date: "19/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Goal


- Recognize when logistic regression can be used, 

    - dependent variable is dichotomous or binary
    
    - predicator (independent) variables can be continuous or categorical. They are related to the probability or odds of the outcome variable. 


#### Example

Suppose that we are interested in the factors that influence whether a ESBL E.Coli is present or not in food samples. The outcome (response) variable is binary (0/1); no present or present. The predictor variables of interest are the origin of the food and the type of food.





### Topics


- What is logistic regression?


- Conducting logistic regression analysis


- Interpret significance testing 

- Appraise whether confounding or interactions are present 

- Interpret logistic regression output 



#### What is logistic regression
    
    - The process of fiting a line or a curve is called regression 
    
    - We use regression analysis to estimate relatioships among variables
    
    - In particular: the relationship between a dependent and 1 or more independent variables
  
    - The formal representation of a relationship is an __statistical model__
    
    - $$p(X)=\beta_0+\beta_1X$$
  
    - The problem with this approach is that when fiting a straight line for a binary dependent variable the probability $p(x)$ for some values of $X$ can go above 1 or below 0. See the following Figure. The probability of Y goes below zero to above one. 
  
  
```{r lin, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

data(mtcars)
dat <- subset(mtcars, select=c(mpg, am, vs))

ggplot(dat, aes(x=mpg, y=vs)) + geom_point() + 
  #stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) +
  #geom_line() #+
  geom_smooth(method = lm, formula = y ~ x) + labs(x = "X", y = "Y")
```

    - to avoid this, we can use the logistic function to model 
    
  $p(X)$ 
  
      as follows:
  
  
  $$\frac{p(X)}{1-p(X)}=e^{\beta_0+\beta_1X}$$


    - where 
    
  $$\frac{p(X)}{1-p(X)}$$ are the odds ratio
    
    
    - if we bring the log we get
    
    
  $$log(\frac{p(X)}1-p{X})=\beta_0+\beta_1X$$
    
    
And we have our probability between 0 and 1.


    
```{r echo=FALSE, message=FALSE, warning=FALSE}
data(mtcars)
dat <- subset(mtcars, select=c(mpg, am, vs))
ggplot(dat, aes(x=mpg, y=vs)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) +
  labs(x = "X", y = "Y")
  #geom_line() #+
  #geom_smooth(method = lm, formula = y ~ x) + labs(x = NULL, y = NULL)
#http://wiki.cge.dtu.dk:10090/display/CWP/1.+Sequencing+projects%3A+getting+started
#user: cge
#password: cge_master
#our* wiki :)

```

    - a logistic regression models the probability that its outcome belongs to a certain category

- Model building strategy: 
  
        a) univariate analysis 


        b) multivariate analysis

  
        c) refitting the model


        d) goodness of fit


### The need for logistic regression


### Antibiotics

```{r}
library(tidyverse)
 library(AMR)
library(help = "AMR")
data("antibiotics")
an <- antibiotics %>%
  select(group, name, ab, atc_group1, atc_group2)
pen <- an%>%
  filter(group == "penicillins")
unique(an$group)
```

```{r}
library(here)
frm <- read_csv(here("data", "farmN.csv"))
names(frm)
unique(frm$ast)
frm$ast <- with(frm, recode(ast, Sensitive = "0", Resistance = "1"))
unique(frm$ast)
unique(frm$btype)
frm$btype <- with(frm, recode(btype, Layers = "layers"))
unique(frm$btype)
lr <- glm(as.factor(ast) ~ btype, data = frm, family = binomial)
summary(lr) # significant
unique(frm$ftype)
frm$ftype <- with(frm, recode(ftype, "Semi-commercial" = "small", Backyard = "small", 
                               "Small scale" = "small", "Large scale" = "large", "Large Scale" = "large",
                               "Medium scale" = "large"))
lr <- glm(as.factor(ast) ~ ftype, data = frm, family = binomial)
summary(lr)
unique(frm$loc)
lr <- glm(as.factor(ast) ~ loc, data = frm, family = binomial)
summary(lr)
lr <- glm(as.factor(ast) ~ numB, data = frm, family = binomial)
summary(lr)
lr <- glm(as.factor(ast) ~ usage, data = frm, family = binomial)
summary(lr)
lr <- glm(as.factor(ast) ~ period, data = frm, family = binomial)
summary(lr)


lr <- glm(as.factor(ast) ~ btype + ftype + numB + usage, data = frm, family = binomial)
summary(lr)


lr <- glm(as.factor(ast) ~ btype + ftype + numB, data = frm, family = binomial)
summary(lr)


lr <- glm(as.factor(ast) ~ btype + numB, data = frm, family = binomial)
summary(lr)

exp(-2.082)
library(rms)
ddist <- with(frm, datadist(btype, numB, usage, ftype))
options(datadist='ddist')
b <- lrm(as.factor(ast) ~ btype + ftype + numB + usage, data = frm, x=TRUE, y=TRUE)
fastbw(b)


nig <- read_csv(here("data", "nig2.csv"))

names(nig)

names(nig)[5] <- "ageg"
unique(nig$ageg)
nig$ageg <- with(nig, recode(ageg, "Layers" = "layers", "Broiler Starter" = "broiler", "Broiler Finisher" = "broiler", "Spent Layers" = "layers",  "Growers" = "broiler", "Chicks" = "broiler"))


ft <- glm(as.factor(Molecular) ~  Livestock, data = nig, family = binomial)
summary(ft)
ft <- glm(as.factor(Molecular) ~  Flock_Size, data = nig, family = binomial)
summary(ft)
ft <- glm(as.factor(Molecular) ~  Production_System, data = nig, family = binomial)
summary(ft)
ft <- glm(as.factor(Molecular) ~  Waste_management, data = nig, family = binomial)
summary(ft)
ft <- glm(as.factor(Molecular) ~  ageg, data = nig, family = binomial)
summary(ft)
ft <- glm(as.factor(Molecular) ~  Foot_dip, data = nig, family = binomial)
summary(ft)
bl <- lrm(as.factor(Molecular) ~ Foot_dip + Waste_management + Production_System + Livestock, data = nig)#, x=TRUE, y=TRUE)
fastbw(bl)
```





